Semi-inclusive deep-inelastic scattering in Wandzura--Wilczek-type approximation

### 2019-04-01 arxiv2/test-WW-type-v2F.tex
AP

%\newcommand{\AP}[1]{\red{#1}}
\newcommand{\ak}[1]{\orange{#1}}
\newcommand{\Bernie}[1]{{\color[rgb]{0,0.5,0.75}#1}}
\newcommand{\ps}[1]{\blue{#1}}
\newcommand{\gs}[1]{{\color[rgb]{0.65,0,0.65}#1}}
%\newcommand{\bp}[1]{{\color[rgb]{0,0.65,0.35}#1}}
%\newcommand{\ms}[1]{\darkgreen{#1}}
%\newcommand{\kt}[1]{{\color[rgb]{0.1,0.5,0}#1}}

@@

At subleading order in the $1/Q$ expansion one has
\ba\hspace{-1cm}
   &&   \frac{d^6\sigma_{\rm subleading}}{dx\,dy\,dz\,d\psi_l\,d\phi_h\,d \Phperp^2}
   =	\frac{1}{4\pi} \; \frac{d \hat{\sigma}}{dy} \; F_{UU}(x,z,\Phperp^2)
        \Biggl\{
          \cos(\phi_h)\,p_3\,A_{UU}^{\cos(\phi_h)}
	\nonumber\\
   && \hspace{2cm}
	+ \lambda\sin(\phi_h)\,p_4\,A_{LU}^{\sin(\phi_h)}
	+ S_L\sin(\phi_h)\,p_3\,A_{UL}^{\sin(\phi_h)}
	+ \lambda\,S_L\cos(\phi_h)\,p_4\,A_{LL}^{\cos(\phi_h)}\phantom{\frac11}
	\nonumber\\
   && \hspace{2cm}
	+ S_T\sin(2\phi_h-\phi_S)\,p_3\,A_{UT}^{\sin(2\phi_h-\phi_S)}
        + S_T\sin(\phi_S)\,p_3\,A_{UT}^{\sin(\phi_S)} \phantom{\frac11}\nonumber\\
   && \hspace{2cm}
  	+ \lambda\,S_T\cos(\phi_S)\,p_4\,A_{LT}^{\cos(\phi_S)}
        + \lambda\,S_T\cos(2\phi_h-\phi_S)\,p_4\,A_{LT}^{\cos(2\phi_h-\phi_S)}
	\Biggr\} \, .
   \hspace{1cm} \label{Eq:SIDIS-subleading}
\ea\end{subequations}
Neglecting $1/Q^2$ corrections, the kinematic prefactors $p_i$ are given by
% \ba\label{Eq:y-prefactors}
% &&	p_1 = \varepsilon \equiv
%        \frac{1-y -\frac{1}{4}\slim \gamma^2 y^2}{1-y
%        + \frac{1}{2}\slim y^2 +\frac{1}{4}\slim \gamma^2 y^2}
%        \approx \frac{1-y}{1-y+\frac12\,y^2} 		\;,\nonumber\\
% &&	p_2 = \sqrt{1-\epsilon^2} \approx
%        \frac{y(1-\frac12\,y)}{1-y+\frac12\,y^2} 	\;,\nonumber\\
% &&	p_3 =  \sqrt{2\varepsilon(1+\varepsilon)}
%        \approx \frac{(2-y)\sqrt{1-y}}{1-y+\frac12\,y^2}\;,\nonumber\\
% &&	p_4 =  \sqrt{2\varepsilon(1-\varepsilon)}
%        \approx \frac{y\sqrt{1-y}}{1-y+\frac12\,y^2} 	\;,\;\;\;
% \ea
\be\label{Eq:y-prefactors}
	p_1 = \frac{1-y}{1-y+\frac12\,y^2} 		\, , \;\;\;
	p_2 = \frac{y(1-\frac12\,y)}{1-y+\frac12\,y^2}	\, , \;\;\;
	p_3 = \frac{(2-y)\sqrt{1-y}}{1-y+\frac12\,y^2} 	\, , \;\;\;
	p_4 = \frac{y\sqrt{1-y}}{1-y+\frac12\,y^2}     	\, ,
\ee
and the asymmetries {$A_{XY}^{\rm weight}$,  are defined in terms of
structure functions $F_{XY}^{\rm weight}$, as follows}
\be
	A_{XY}^{\rm weight}\equiv A_{XY}^{\rm weight}(x,z,\Phperp)=
	\frac{F_{XY}^{\rm weight}(x,z,\Phperp)}{F_{UU}(x,z,\Phperp)}.
\ee
Here, the first subscript $X=U(L)$ denotes the unpolarized beam
(longitudinally polarized beam with helicity $\lambda$). The second
subscript $Y=U(L\text{ or }T)$ refers to the target, which can be unpolarized
(longitudinally or transversely polarized with respect to the virtual photon).
The superscript ``weight'' indicates the azimuthal dependence with no index
indicating an isotropic angular distribution of the produced hadrons.

@@

\ak{As it was mentioned in the Introduction}
\ps{one important limitation concerns the fact that the WW-type approximations
are not preserved under $Q^2$ evolution. \ak{Still} some intuition can be 
obtained from the collinear case: the evolution equations for $g_T^a(x)$ 
and $h_L^a(x)$ exhibit complicated mixing patterns typical for higher 
twist functions, which simplify to DGLAP-type evolutions in the limit 
of a large number of colors $N_c$ and in the limit of large-$x$ 
\cite{Ali:1991em,Koike:1994st,Balitsky:1996uh,Belitsky:1997zw}. 
These evolution equations differ from those of the leading-twist
functions $g_1^a(x)$ and $h_1^a(x)$. This point is of not much
practical relevance here, because the $Q^2$ varies only moderately
between JLab, HERMES, and COMPASS (except for the largest $x$-bins).
However, if for some reason the $\bar{q}gq$-terms were found to be
very small at one renormalization scale, it is not guaranteed the
WW-type approximation will work equally well also at other scales. 
 More theoretical work is required, especially also in order to 
understand $k_\perp$-evolution effects at subleading twist.}

@@

The asymmetries $A_{UT}^{\sin(\phi_h-\phi_S)}= F_{UT}^{\sin(\phi_h-\phi_S)}/F_{UU}$
obtained from the fit \cite{Anselmino:2011gs}
are plotted in Fig.~\ref{aut_f1t_jlab} as functions of $x$ in comparison
to HERMES \cite{Airapetian:2009ae} and COMPASS \cite{Adolph:2012sp} data
on respectively charged pion and hadron production from a proton target.
	Notice that the COMPASS data points seem to be below the
	theoretical curves which may indicate evolution
	effects~\cite{Aybat:2011ta,Anselmino:2012aa}.
We do not show here the description of the $P_{hT}$--dependence of the
data but it is well described by the fit of \cite{Anselmino:2011gs}
which confirms that the Gaussian model works also in this case.

@@

The $F_{UT}^{\sin(\phi_h+\phi_S)}$ structure function of the SIDIS cross section is
due to the convolution of the transversity distribution $h_1$ and the Collins
FF $H_1^\perp$. Transversity can in principle be accessed also as a PDF in
Drell--Yan or in dihadron production
\cite{Bacchetta:2002ux,Bacchetta:2003vn,Bacchetta:2011ip,Bacchetta:2012ty,
Radici:2015mwa,Radici:2018iag}. The first polarized Drell--Yan data were
measured by COMPASS~\cite{Aghasyan:2017jop,Parsamyan:2018zju}.
It describes the distribution of transversely polarized quarks
in a transversely polarized nucleon, and is the only source of information
on the tensor charge of the nucleon. The Collins FF $H_1^\perp$ decodes the
fundamental correlation between the transverse spin of a fragmenting quark
and the transverse momentum of the final hadron produced~\cite{Collins:1992kk}.

There are many extractions of $h_1$ and $H_1^\perp$ from a
combined fit of SIDIS and $e^+e^-$ data, for instance those of
Refs.~\cite{Anselmino:2013vqa,Kang:2014zza,Anselmino:2015sxa}.
In this work we will use the extractions of $h_1$ and $H_1^\perp$
from Ref.~\cite{Anselmino:2013vqa}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
\centering
\includegraphics[width=0.45\textwidth]{\FigPath/AUTCollins_x.pdf}
{\tiny(a)}
\includegraphics[width=0.45\textwidth]{\FigPath/AUTCollins_COMPASS_x.pdf}
{\tiny(b)}
\caption{\label{aut_h1_jlab}  Collins asymmetry
	$A_{UT,  \langle y\rangle}^{\sin(\phi_h+\phi_S)}$ for a proton target 
	as function of $ x $
	based on the fit \cite{Anselmino:2013vqa} in comparison to
	\ps{(a)} HERMES \cite{Airapetian:2010ds} and
	\ps{(b)} COMPASS \cite{Adolph:2014zba} data [where
	$(-1)A_{UT}^{\sin(\phi_h+\phi_S)}$ is shown since COMPASS 
	\gs{uses an opposite sign in its Collins asymmetry convention}].}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%\newcommand{\AP}[1]{\red{#1}}
\newcommand{\ak}[1]{\orange{#1}}
\newcommand{\Bernie}[1]{{\color[rgb]{0,0.5,0.75}#1}}
\newcommand{\ps}[1]{\blue{#1}}
\newcommand{\gs}[1]{{\color[rgb]{0.65,0,0.65}#1}}
\newcommand{\bp}[1]{{\color[rgb]{0,0.65,0.35}#1}}
%\newcommand{\ms}[1]{\darkgreen{#1}}
%\newcommand{\kt}[1]{{\color[rgb]{0.1,0.5,0}#1}}

@@

At subleading order in the $1/Q$ expansion one has
\ba\hspace{-1cm}
   &&   \frac{d^6\sigma_{\rm subleading}}{dx\,dy\,dz\,d\psi_l\,d\phi_h\,d \Phperp^2}
   =	\frac{1}{4\pi} \; \frac{d \hat{\sigma}}{dy} \; F_{UU}(x,z,\Phperp^2)
        \Biggl\{
          \cos(\phi_h)\,p_3\,A_{UU}^{\cos(\phi_h)}
	\nonumber\\
   && \hspace{2cm}
	+ \lambda\sin(\phi_h)\,p_4\,A_{LU}^{\sin(\phi_h)}
	+ S_L\sin(\phi_h)\,p_3\,A_{UL}^{\sin(\phi_h)}
	+ \lambda\,S_L\cos(\phi_h)\,p_4\,A_{LL}^{\cos(\phi_h)}\phantom{\frac11}
	\nonumber\\
   && \hspace{2cm}
	+ S_T\sin(2\phi_h-\phi_S)\,p_3\,A_{UT}^{\sin(2\phi_h-\phi_S)}
        + S_T\sin(\phi_S)\,p_3\,A_{UT}^{\sin(\phi_S)} \phantom{\frac11}\nonumber\\
   && \hspace{2cm}
  	+ \lambda\,S_T\cos(\phi_S)\,p_4\,A_{LT}^{\cos(\phi_S)}
        + \lambda\,S_T\cos(2\phi_h-\phi_S)\,p_4\,A_{LT}^{\cos(2\phi_h-\phi_S)}
	\Biggr\} \, .
   \hspace{1cm} \label{Eq:SIDIS-subleading}
\ea\end{subequations}
Neglecting $1/Q^2$ corrections, the kinematic prefactors $p_i$ are given by
% \ba\label{Eq:y-prefactors}
% &&	p_1 = \varepsilon \equiv
%        \frac{1-y -\frac{1}{4}\slim \gamma^2 y^2}{1-y
%        + \frac{1}{2}\slim y^2 +\frac{1}{4}\slim \gamma^2 y^2}
%        \approx \frac{1-y}{1-y+\frac12\,y^2} 		\;,\nonumber\\
% &&	p_2 = \sqrt{1-\epsilon^2} \approx
%        \frac{y(1-\frac12\,y)}{1-y+\frac12\,y^2} 	\;,\nonumber\\
% &&	p_3 =  \sqrt{2\varepsilon(1+\varepsilon)}
%        \approx \frac{(2-y)\sqrt{1-y}}{1-y+\frac12\,y^2}\;,\nonumber\\
% &&	p_4 =  \sqrt{2\varepsilon(1-\varepsilon)}
%        \approx \frac{y\sqrt{1-y}}{1-y+\frac12\,y^2} 	\;,\;\;\;
% \ea
\be\label{Eq:y-prefactors}
	p_1 = \frac{1-y}{1-y+\frac12\,y^2} 		\, , \;\;\;
	p_2 = \frac{y(1-\frac12\,y)}{1-y+\frac12\,y^2}	\, , \;\;\;
	p_3 = \frac{(2-y)\sqrt{1-y}}{1-y+\frac12\,y^2} 	\, , \;\;\;
	p_4 = \frac{y\sqrt{1-y}}{1-y+\frac12\,y^2}     	\, ,
\ee
and the asymmetries {$A_{XY}^{\rm weight}$,  are defined in terms of
structure functions $F_{XY}^{\rm weight}$, as follows}
\be
	A_{XY}^{\rm weight}\equiv A_{XY}^{\rm weight}(x,z,\Phperp)=
	\frac{F_{XY}^{\rm weight}(x,z,\Phperp)}{F_{UU}(x,z,\Phperp)}.
\ee
Here, the first subscript $X=U(L)$ denotes the unpolarized beam
(longitudinally polarized beam with helicity $\lambda$). The second
subscript $Y=U(L\text{ or }T)$ refers to the target, which can be unpolarized
(longitudinally or transversely polarized with respect to the virtual photon).
The superscript ``weight'' indicates the azimuthal dependence with no index
indicating \bp{a $\phi$-independent} distribution of the produced hadrons.

@@

\ak{As it was mentioned in the Introduction}
\ps{one important limitation concerns the fact that the WW-type approximations
are not preserved under $Q^2$ evolution. \ak{Still} some intuition can be 
obtained from the collinear case: the evolution equations for $g_T^a(x)$ 
and $h_L^a(x)$ exhibit complicated mixing patterns typical for higher 
twist functions, which simplify to DGLAP-type evolutions in the limit 
of a large number of colors $N_c$ and in the limit of large-$x$ 
\cite{Ali:1991em,Koike:1994st,Balitsky:1996uh,Belitsky:1997zw}. 
These evolution equations differ from those of the leading-twist
functions $g_1^a(x)$ and $h_1^a(x)$. 
% This point is of not much
% practical relevance here, because the $Q^2$ varies only moderately
% between JLab, HERMES, and COMPASS (except for the largest $x$-bins).
% However, if for some reason the $\bar{q}gq$-terms were found to be
% very small at one renormalization scale, it is not guaranteed the
% WW-type approximation will work equally well also at other scales. 
% More theoretical work is required, especially also in order to 
% understand $k_\perp$-evolution effects at subleading twist.
\bp{However, since $Q^2$ varies moderately in the considered experiments 
(e.g.\ for common values of $x$ the $Q^2$ at COMPASS is only about factor 2 
larger than at HERMES), this point is not a major uncertainty in our study.
More theoretical work will be required to understand $k_\perp$-evolution 
effects of subleading twist TMDs in future experiments (EIC) 
covering larger $Q^2$ ranges.}}

@@

The asymmetries $A_{UT}^{\sin(\phi_h-\phi_S)}= F_{UT}^{\sin(\phi_h-\phi_S)}/F_{UU}$
obtained from the fit \cite{Anselmino:2011gs}
are plotted in Fig.~\ref{aut_f1t_jlab} as functions of $x$ in comparison
to HERMES \cite{Airapetian:2009ae} and COMPASS \cite{Adolph:2012sp} data
on respectively charged pion and hadron production from a proton target.
	Notice that the COMPASS data points seem to be below the
	theoretical curves which may indicate evolution
	effects~\cite{Aybat:2011ta,Anselmino:2012aa}.
We do not show here the description of the $P_{hT}$--dependence of the
data but it is well described by the fit of \cite{Anselmino:2011gs}
which confirms that the Gaussian model works also in this case.
\bp{The Sivers function is predicted to enter the description of hadron-hadron 
collisions (with transversely polarized protons) with an opposite sign 
compared to SIDIS \cite{Collins:2002kn,Brodsky:2002cx,Brodsky:2002rv}.
Recent results on single-spin asymmetries in weak boson production 
from RHIC \cite{Adamczyk:2015gyk} and Drell--Yan from COMPASS 
\cite{Aghasyan:2017jop,Parsamyan:2018zju} are consistent 
with this prediction.}

@@

The $F_{UT}^{\sin(\phi_h+\phi_S)}$ structure function of the SIDIS cross section is
due to the convolution of the transversity distribution $h_1$ and the Collins
FF $H_1^\perp$. 
It describes the distribution of transversely polarized quarks 
in a transversely polarized nucleon, and is the only source of information
on the tensor charge of the nucleon. 
\bp{Transversity can also be accessed as a PDF in Drell--Yan or dihadron 
production \cite{Bacchetta:2002ux,Bacchetta:2003vn,Bacchetta:2011ip,
Bacchetta:2012ty,Radici:2015mwa,Radici:2018iag}.} 
% The first polarized Drell--Yan data were measured by 
% COMPASS~\cite{Aghasyan:2017jop,Parsamyan:2018zju}.
The Collins FF $H_1^\perp$ decodes the
fundamental correlation between the transverse spin of a fragmenting quark
and the transverse momentum of the produced final hadron~\cite{Collins:1992kk}.
%
There are many extractions of $h_1$ and $H_1^\perp$ from 
combined fits of SIDIS and $e^+e^-$ data, for instance those of
Refs.~\cite{Anselmino:2013vqa,Kang:2014zza,Anselmino:2015sxa}.
In this work we will use the extractions of $h_1$ and $H_1^\perp$
from Ref.~\cite{Anselmino:2013vqa}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
\centering
\includegraphics[width=0.45\textwidth]{\FigPath/AUTCollins_x.pdf}
{\tiny(a)}
\includegraphics[width=0.45\textwidth]{\FigPath/AUTCollins_COMPASS_x.pdf}
{\tiny(b)}
\caption{\label{aut_h1_jlab}  Collins asymmetry
	$A_{UT,  \langle y\rangle}^{\sin(\phi_h+\phi_S)}$ for a proton target 
	as function of $ x $
	based on the fit \cite{Anselmino:2013vqa} in comparison to
	\ps{(a)} HERMES \cite{Airapetian:2010ds} and
	\ps{(b)} COMPASS \cite{Adolph:2014zba} data. Notice that we multiplied COMPASS data by a factor $-1$ in order to match our convention of the Collins angle.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
